### 2024-06-15 ms.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{setspace}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={A general model for the evolution of nuptial gift-giving},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{amsmath}
\usepackage{natbib}
\usepackage{lineno}
\usepackage{caption}
\usepackage[utf8]{inputenc}
\bibliographystyle{amnatnat}
\linenumbers
\modulolinenumbers[2]
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi

\title{A general model for the evolution of nuptial gift-giving}
\author{Anders P. Charmouh\(^{1,4, 5}\), Trine Bilde\(^{2}\), Greta
Bocedi\(^{1}\), A. Bradley Duthie\(^{3}\)}
\date{{[}1{]} School of Biological Sciences, University of Aberdeen, UK.
{[}2{]} Department of Biology, Aarhus University, Denmark. {[}3{]}
Biological and Environmental Sciences, University of Stirling, UK.
{[}4{]} Bioinformatics Research Centre Aarhus University, University
City 81, building 1872, 3rd floor, DK-8000 Aarhus C, Denmark {[}5{]}
Email: \href{mailto:apc@birc.au.dk}{\nolinkurl{apc@birc.au.dk}}}

\begin{document}
\maketitle

\setstretch{2}
\textbf{Key words:} Nuptial gifts, male search, mate interactions,
modelling, \emph{Pisaura}

\hypertarget{abstract}{%
\section{Abstract}\label{abstract}}

Nuptial gift-giving occurs in several taxonomic groups including
insects, snails, birds, squid, arachnids and humans. Although this trait
has evolved many times independently, no general framework has been
developed to predict the conditions necessary for nuptial gift-giving to
evolve. We use a time-in time-out model to derive analytical results
describing the requirements necessary for selection to favour nuptial
gift-giving. Specifically, selection will favour nuptial gift-giving if
the fitness increase caused by gift-giving exceeds the product of
expected gift search time and encounter rate of the opposite sex.
Selection will favour choosiness in the opposite sex if the value of a
nuptial gift exceeds the inverse of the time taken to produce offspring
multiplied by the rate at which mates with nuptial gifts are
encountered. Importantly, selection can differ between the sexes,
potentially causing sexual conflict. We test these results using an
individual-based model applied to a system of nuptial gift-giving
spiders, \emph{Pisaura mirabilis}, by estimating parameter values using
experimental data from several studies. Our results provide a general
framework for understanding when the evolution of nuptial gift-giving
can occur and provide novel insight into the evolution of worthless
nuptial gifts, occurring in multiple taxonomic groups with implications
for understanding parental investment.

\hypertarget{lay-summary}{%
\section{Lay summary}\label{lay-summary}}

If you have ever bought a gift to impress a prospective partner, you are
not alone. Such behaviour is not unique to humans, but giving gifts
during courtship is common in many different animals, including some
species of insects, snails, birds, squid and spiders. In animals, such
gifts can be costly, but offering them may be worth it from an
evolutionary perspective if they increase the reproductive success of
the individual who provides them. While using gifts during courtship
(often called nuptial gifts) is widespread among animals, it is not
universal. Why have some groups of animals evolved to engage in nuptial
gift-giving while others have not? We explore this question by building
a mathematical model, using computer simulations, and using data from a
nuptial gift-giving species of spider to see how receiving nuptial gifts
might affect reproductive success. We find conditions under which
selection should favour males to provide nuptial gifts and females
should exhibit a preference for gifts. These conditions explain why some
species engage in nuptial gift-giving while others do not.
Interestingly, we find that these inequalities differ between males and
female, illustrating that the evolutionary interest of the two sexes do
not always overlap.

\hypertarget{teaser-text}{%
\section{Teaser text}\label{teaser-text}}

The results presented provide a general explanation of the evolution of
nuptial gift-giving which is a behaviour pattern that has evolved many
times independently in diverse parts of the animal kingdom. The
explanation is general but at the same time quantitatively precise, thus
giving a rigorous answer to the conundrum of why nuptial gift-giving
occurs in some groups of animals but not in others. Since research on
nuptial gift-giving has, and continues to, inspire an enormous amount of
work, the results presented here are highly relevant to an active field
of both theoretical and empirical research.

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Nuptial gift-giving occurs when the choosy sex (usually the female)
receives gifts from the opposite sex (usually the male) during
courtship. It is a widespread phenomenon, occurring within several
diverse taxonomic groups such as insects, snails, birds, squid,
arachnids and humans (\protect\hyperlink{ref-Lewis2012}{Lewis \& South,
2012}; \protect\hyperlink{ref-Albo2014}{Albo \emph{et al.}, 2014};
\protect\hyperlink{ref-Lewis2014}{Lewis \emph{et al.}, 2014}). Despite
the ubiquity of this behaviour, little effort has been made to
conceptualise the evolution of nuptial gift-giving within a general
modelling framework (\protect\hyperlink{ref-Lewis2014}{Lewis \emph{et
al.}, 2014}; \protect\hyperlink{ref-Iwasa2022}{Iwasa \& Yamaguchi,
2022}). Recent models describing the evolution of nuptial gift-giving
have focused on co-evolution between male nuptial gift-giving and female
propensity to remate, and evolutionarily stable nuptial gift sizes
(\protect\hyperlink{ref-Kamimura2021}{Kamimura \emph{et al.}, 2021};
\protect\hyperlink{ref-Iwasa2022}{Iwasa \& Yamaguchi, 2022}), but a
general framework describing the conditions necessary for nuptial
gift-giving to be initially favoured by selection is needed to
understand when gift-giving should evolve.

Nuptial gift-giving may allow males to increase fitness by acquiring
additional mates, indirect benefits (by increasing the fitness of
offspring), prolonged copulations, and success in sperm competition
(\protect\hyperlink{ref-Albo2013}{Albo \emph{et al.}, 2013};
\protect\hyperlink{ref-Ghislandi2014}{Ghislandi \emph{et al.}, 2014};
\protect\hyperlink{ref-Lewis2014}{Lewis \emph{et al.}, 2014}). However,
this potential fitness increase comes at the expense of producing a
nuptial gift, which may be costly in terms of time and resources.
Females may increase their fitness by receiving nutritionally valuable
nuptial gifts, but expressing a preference for males with gifts might
result in a mating opportunity cost if available males without gifts are
rejected. With respect to nuptial gift-giving, the evolutionary
interests of both sexes may not always fully overlap. This can cause
sexual conflict, which is a difference in the evolutionary interest
between sexes that occurs when interaction between sexes results in a
situation where neither sex can achieve an optimal outcome
(\protect\hyperlink{ref-Parker2006}{Parker, 2006}). That is, under some
conditions, it might for example be optimal for males but not females to
mate if females do not benefit from mating with males without nuptial
gifts.

Much work has sought to explain how gift-giving tactics are maintained,
with explanations including condition-dependent strategies, gift-giving
as a way to decrease female aggression during copulation, or gifts as
sensory traps (\protect\hyperlink{ref-Lubin2007}{Lubin \& Bilde, 2007};
\protect\hyperlink{ref-Toft2016}{Toft \& Albo, 2016};
\protect\hyperlink{ref-Ghislandi2018}{Ghislandi \emph{et al.}, 2018};
\protect\hyperlink{ref-Albo2019}{Albo \emph{et al.}, 2019}). An example
of such a system is the nuptial gift-giving nursery-web spider
\emph{Pisaura mirabilis} where males may court females with or without
nuptial gifts (\protect\hyperlink{ref-Bristowe1926}{Bristowe \& Locket,
1926}; \protect\hyperlink{ref-Tuni2013a}{Tuni \emph{et al.}, 2013}).
Here, males may provide females with costly nuptial gifts in the form of
captured arthropod prey, and females may exhibit preference for males
with a nuptial gift by rejecting males without a nuptial gift
(\protect\hyperlink{ref-Albo2013}{Albo \emph{et al.}, 2013}).

We develop a general framework for investigating the evolution of
nuptial gift-giving and choosiness using a time-in, time-out modelling
approach and an individual-based model
(\protect\hyperlink{ref-Clutton-Brock1992}{Clutton-Brock \& Parker,
1992}). Specifically, we derive conditions under which selection will
favour male search for nuptial gifts and female rejection of gift-less
males. We show that selection for searching and choosiness depends on
whether a threshold fitness value of the nuptial gift is exceeded. Our
model demonstrates the importance of nuptial gift cost, sex ratio, and
mate encounter rate in determining the threshold above which selection
will favour the evolution of nuptial gift-giving. Importantly, we show
that the threshold value differs for males and females. We test
predictions of our analytical model by formulating an individual-based
model, which further supports the main theoretical results of our
analytical model. Further, we apply our model to an example system with
nuptial gifts, the nursery web spider \emph{Pisaura mirabilis}, where we
use experimental data to estimate a key model parameter. Our results
provide a general framework for understanding why nuptial gift-giving
evolves in some systems and not in others, how the evolution of nuptial
gift-giving can give rise to sexual conflict, and it provides insight
into the evolution of worthless and deceitful nuptial gifts, which occur
in several different taxonomic (\protect\hyperlink{ref-LeBas2005}{LeBas
\& Hockham, 2005}; \protect\hyperlink{ref-Ghislandi2014}{Ghislandi
\emph{et al.}, 2014}).

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{analytical-model}{%
\subsection{Analytical model}\label{analytical-model}}

We use a time-in and time-out model
(\protect\hyperlink{ref-Clutton-Brock1992}{Clutton-Brock \& Parker,
1992}; \protect\hyperlink{ref-Kokko2001}{Kokko \& Monaghan, 2001};
\protect\hyperlink{ref-Kokko2006}{Kokko \& Ots, 2006}) in which choosy
(female) and non-choosy (male) individuals spend some period of time
within the mating pool searching for a mate (time-in) followed by a
period outside the mating pool (time-out). During time-out, females
spend some duration of time (\(T_{\mathrm{f}}\)) gestating or rearing
(hereafter `processing') offspring. We define the number of offspring
produced by a female per reproductive cycle as \(\lambda\). Since
females enter time-out after mating, this assumption is equivalent to
assuming a system with sequential polyandry. For simplicity, we assume
male time to replenish sperm is negligible, but males can spend some
duration of time (\(T_{\mathrm{m}}\)) out of the mating pool searching
for nuptial gifts.

\begin{figure}
\includegraphics[width=1\linewidth]{inst/img/conceptual_figure} \caption{Conceptual figure inspired by Kokko and Ots (2006) illustrating how the modelling framework maps onto an example of a system wherein nuptial gifts are used, here \textit{Pisaura mirabilis}. Males have a probability of obtaining a nuptial gift while in time-out, which will affect their probability of mating while in time-in. They return to the mating pool (time-in) at a rate determined by the time spent searching for a nuptial gifts ($T_{\mathrm{m}}$) and leave the mating pool (i.e. enter time-out) following the female encounter rate, which is dependent on the ratio of males to females ($\beta$) and the encounter rate ($R$). The choosy sex (females) enter the mating pool at a rate depending on the time spent processing offspring ($T_{\mathrm{f}}$) and leave the mating pool (i.e. enter time-out) at a rate that is dependent on $\beta$ and $R$. Males and females undergo sex-specific mortality $\mu$ during time-in and time-out. Image left to right: (1) male \textit{P. mirabilis}. (2) male \textit{P. mirabilis} presenting nuptial gift (white) to female. (3) Female \textit{P. mirabilis} protecting offspring. Photos: Alamy.}\label{fig:unnamed-chunk-2}
\end{figure}

\hypertarget{criteria-for-male-search-and-female-choosiness}{%
\subsection{Criteria for male search and female
choosiness}\label{criteria-for-male-search-and-female-choosiness}}

The probability \(G\) that a male succeeds in securing a nuptial gift is
defined by,

\begin{equation}
G = 1 - e^{-\frac{1}{\alpha}T_{\mathrm{m}}}.
\end{equation}

In Eq. 1, \(\alpha\) defines the expected search time before
encountering a nuptial gift. Thus, the probability of finding a nuptial
gift is higher the more time \(T_{\mathrm{m}}\) is spent searching.
During time-in, individuals encounter conspecifics at a rate of \(R\). A
focal individual will therefore encounter conspecifics of the opposite
sex at a rate of \(R/2\) if the ratio of males to females in the mating
pool (\(\beta\)) is equal. More generally, males will be encountered at
a rate of \(R\beta/(\beta+1)\), and females will be encountered at a
rate of \(R/(\beta+1)\). An example of how the structure of the time-in
time-out model applies to a system with nuptial gift-giving is given in
Figure 1. We assume that mating with a nuptial gift increases the
fitness of each offspring by an increment of \(\gamma\). We proceed to
find the thresholds \(\gamma_{\mathrm{m}}\) and \(\gamma_{\mathrm{f}}\)
above which males and females are favoured by selection to search for
mates with nuptial gifts and exhibit choosiness for nuptial gifts,
respectively. We show (see Methods) that the initial threshold value of
\(\gamma\) (\(\gamma_{\mathrm{m}}\)) necessary for males to increase
their fitness by investing time searching for a nuptial gift (time that
could otherwise be spent searching for a mate) is,

\begin{equation}
\gamma_{\mathrm{m}} > \alpha \frac{R}{\beta + 1}.
\end{equation}

Inequality 2 means that if nuptial gifts are not abundant and thus
require a long time to find (i.e., high \(\alpha\)), or if males
encounter many females per unit time (i.e., high \(R / (1+\beta)\)),
then the nuptial gift must result in a high fitness increment for
selection to favour gift-searching. In general, when ineq. 2 is
satisfied, we predict selection to favour the evolution of nuptial
gift-giving.

We can similarly predict the conditions for which there is selection for
female choosiness. If \(\gamma\) is sufficiently high, then females
increase their fitness by rejecting males without gifts and mating only
with males that provide nuptial gifts. To illustrate, we assume that all
males in a population search for a duration of \(T_{\mathrm{m}}\), in
which case the threshold fitness increment for females
(\(\gamma_{\mathrm{f}}\)) is,

\begin{equation}
\gamma_{\mathrm{f}} > \frac{1}{T_{\mathrm{f}} R \left(\frac{\beta}{\beta + 1}\right) \left(1 - e^{-\frac{1}{\alpha}T_{\mathrm{m}}} \right)}.
\end{equation}

Inequality 3 shows that as offspring processing time
(\(T_{\mathrm{f}}\)), mate encounter rate (\(R\beta / (\beta + 1)\)), or
the probability of a male finding a nuptial gift
(\(1 - \exp(-T_{\mathrm{m}}/\alpha)\) ) decreases, the threshold value
of fitness above which selection will favour choosiness
(\(\gamma_{\mathrm{f}}\)) increases. This can be understood intuitively
by realising that rejecting a prospective male represents an opportunity
cost for the female. This opportunity cost becomes small if many males
with gifts are encountered, hence the appearance of the rate at which
males with gifts are encountered in the denominator. Figure 2 shows how
\(\gamma_{\mathrm{m}}\) and \(\gamma_{\mathrm{f}}\) change with
increasing \(\alpha\). For \(\gamma_{\mathrm{f}}\), we assume that males
search for the expected time required to obtain a nuptial gift
(\(T_{\mathrm{m}} = \alpha\)). Note that \(\beta\) does not have a
closed form solution given \(R\), \(T_{\mathrm{f}}\), and
\(T_{\mathrm{m}}\), so \(\beta\) was calculated using recursion (see
Methods).

\begin{figure}
\centering
\includegraphics{ms_files/figure-latex/unnamed-chunk-3-1.pdf}
\caption{Fitness thresholds above which males increase their fitness by
searching for nuptial gifts (blue lines; Eq. 2) and females increase
their fitness by rejecting males that do not offer gifts (red lines; Eq.
3). Parameter space includes areas in which males do not search for
nuptial gifts and females are not choosy (A), males search but females
are not choosy (B), females would be choosy but males do not search (C),
and males search and females are choosy (D). Arrows in panel a indicate
the effect of increasing interaction rate (\(R\)), female time-out
(\(T_{\mathrm{f}}\)), and male search time (\(T_{\mathrm{m}}\)). As an
example, trajectories for \(T_{\mathrm{f}} = 2\), and
\(T_{\mathrm{m}} = \alpha\) are shown for values of \(R=2\) (panel a)
and \(R = 1\) (panel b). Females are assumed to be the choosy sex, which
is maintained as long as \(\alpha < T_{\mathrm{f}}\).}
\end{figure}

The analytical framework predicts 4 zones, which are delineated by
inequalities 2 and 3 and describe the initial thresholds for favouring
search of nuptial gifts in males and choosiness for nuptial gifts in
females (Figure 2a). Consequently, the modelling framework gives a
description of the conditions under which nuptial gift-giving is
expected to occur (Figure 2a, Zone D) and the conditions under which
only selection for male searching (Figure 2a, Zone B) or female
choosiness (Figure 2a, Zone C) are predicted. These results therefore
highlight the potential for sexual conflict over nuptial gift-giving.

\hypertarget{evolution-of-male-search-and-female-choosiness}{%
\subsection{Evolution of male search and female
choosiness}\label{evolution-of-male-search-and-female-choosiness}}

We used an individual-based model (IBM) to simulate the evolution of
nuptial gift-giving and female choosiness from an ancestral condition in
which neither exists. The IBM was written to satisfy the assumptions of
our analytical time-in and time-out model as much as practical (see
Supporting Information S1). Using the IBM, we modelled a
spatially-implicit, finite population of females and males. At each time
step, some individuals enter or remain within the mating pool (time-in),
where they potentially interact and mate. After mating, males and
females may leave the mating pool to search for nuptial gifts and to
produce offspring, respectively (time-out). Mortality occurs with a
fixed probability in each time step, then a ceiling regulation is
applied to limit population growth (see Methods).

The rates at which females are encountered by males
\(R_{\mathrm{f},\mathrm{m}}\) and males with nuptial gifts
\(R_{\mathrm{m_{G}}, \mathrm{f}}\) are encountered by females are both
calculated directly from the IBM, thereby modelling how these rates
might be estimated from empirical data, so the male threshold for
increasing fitness by searching is,

\begin{equation}
\gamma_{\mathrm{m, IBM}} > \alpha R_{\mathrm{f},\mathrm{m}}.
\end{equation}

Similarly, female threshold for increasing fitness by choosiness is,

\begin{equation}
\gamma_{\mathrm{f, IBM}} > \frac{1}{T_{\mathrm{f}}R_{\mathrm{m_{G}}, \mathrm{f}}}.
\end{equation}

Consequently, the IBM and the analytical model differ slightly (e.g.,
time is discrete in the IBM but continuous the analytical model, and in
the IBM, a fitness increment is applied to the focal female in the form
of birthrate increase rather than offspring fitness; see Supporting
Information S1 for details). But the predicted thresholds are
theoretically equivalent and yield predictions that are qualitatively
the same (Figure 3).

\begin{figure}
\centering
\includegraphics{ms_files/figure-latex/unnamed-chunk-4-1.pdf}
\caption{The coevolution of male search and female choosiness as a
function of nuptial gift search time (\(\alpha\)). Points show where the
lower 95\% confidence interval of female choosiness (red) and male
search (blue) exceeds zero, indicating evolution of choosiness or
nuptial gift search. Each point includes data from 3200 replicate
simulations with identical starting conditions. Red and blue lines show
thresholds above which the mathematical model predicts that females
should be choosy and males should search, respectively (in agreement
with Figure 2). Minor deviations between the analytical model and the
simulation results are expected due to the finite population size, the
substantial stochasticity inherent to the simulation model, and the
potential for coevolution (for details, see Supporting Information S1).
The number of individuals in the population remained at or near carrying
capacity of \(K = 1000\). In each time step, up to 3000 total pair-wise
interactions occurred. Expected female processing time was set to
\(T_{\mathrm{f}}=2\) time steps, and \(\gamma\) and \(\alpha\) values in
the range {[}0.0, 2.0{]} and {[}0.1, 2.0{]}, respectively, were used.}
\end{figure}

Male thresholds \(\gamma_{\mathrm{m, IBM}}\) given by ineq. 4 accurately
predict the evolution of searching in the IBM across \(\alpha\) values,
and the female threshold \(\gamma_{\mathrm{f, IBM}}\) (ineq. 5) predicts
the evolution of female choice (Figure 3). In other words, IBM
simulations demonstrate that nuptial gift search in males, and
choosiness in females, will evolve from an ancestral state of no
searching and no choosiness in similar parameter space (Figure 3) as
predicted by the analytical model (Figure 2b). We further ran the IBM
with realistic values of \(\gamma\), estimated using data from the
\emph{P. mirabilis} system wherein choosiness among females, and nuptial
gift search among males, occur (Figure 4). We found that our IBM
predicts both the evolution of choosiness and nuptial gift searching
observed in the \emph{P. mirabilis} system.

\begin{figure}
\centering
\includegraphics{ms_files/figure-latex/unnamed-chunk-5-1.pdf}
\caption{The joint evolution of male search and female choosiness using
a nuptial gift fitness increment (\(\gamma\)) that was estimated from
experimental data for the species \emph{Pisaura mirabilis} (mean
\(\gamma = 3.29 \pm\) SE via propagation of error for estimates of
nuptial feeding a non-nuptial feeding groups: {[}1.47; 8.32{]}). Points
show where the 95\% confidence interval exceeds 0 for female choosiness
(red) and male search (blue). Each point includes data from 100
replicate simulations with identical starting conditions. The red line
shows the threshold above which females should be choosy and the blue
line shows the threshold above which males should search. As predicted
by the analytical model, both male search and female choosiness evolved
for a range of \(\gamma\) values around the empirical estimates. This
occurs because these values result in a \(\gamma\) above the fitness
threshold necessary for selection to favour male search of nuptial gifts
(blue line) and female choosiness (red line). The number of individuals
in the population remained at or near carrying capacity of \(K = 1000\).
In each time step, up to 3000 total pair-wise interactions occurred.
Female processing time was set to \(T_{\mathrm{f}}=2\) time steps, and
\(\gamma\) and \(\alpha\) values in the {[}0.2, 2.0{]}, respectively,
were used. The parameter \(\gamma\) was estimated as relative increase
in offspring production such that \(\gamma\) is the factor by which
fitness increases (relative to the baseline fitness) given a nuptial
gift (See Supporting Information S5).}
\end{figure}

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

Nuptial gift-giving has arisen several times independently throughout
the animal kingdom (\protect\hyperlink{ref-Lewis2012}{Lewis \& South,
2012}), so understanding how selection favours nuptial gift giving and
choosiness is important for a broad range of mating systems. We provide
a general framework that defines the necessary conditions for selection
to favour the evolution of nuptial gift-giving. We show that males
should give nuptial gifts if the value of a nuptial gift exceeds a
threshold dependent on the encounter rate between males and females and
the cost or time necessary to find or produce a nuptial gift (see ineq.
2). This result makes intuitive sense because if males rarely encounter
females, time searching for a gift is a minor cost relative to mate
search time. If males encounter many females, it is not worth seeking
nuptial gifts unless gifts are very valuable since the male will meet
many prospective mates, and nuptial gift search time might come at a
cost of decreased mating opportunities. In practice, male biased sex
ratios will not necessarily favour male search for nuptial gifts if the
female encounter rate is very high, so the key variable is how often
males and females encounter each other. If the search time or cost of
finding a nuptial gift is high, nuptial gifts must be very valuable
before search is favoured by selection.

\hypertarget{threshold-fitness-values}{%
\subsection{Threshold fitness values}\label{threshold-fitness-values}}

Importantly, we show that the threshold nuptial gift value at which
females are favoured to express choosiness for nuptial gifts is rarely
equivalent to the threshold value at which males are favoured to search
for nuptial gifts, potentially leading to sexual conflict
(\protect\hyperlink{ref-Arnqvist2005a}{Arnqvist \& Rowe, 2005};
\protect\hyperlink{ref-Oliveira2008}{Oliveira \emph{et al.}, 2008}).
Here, we are defining sexual conflict as occurring when interactions
between sexes result in situation where both sexes cannot achieve an
optimal outcome simultaneously
(\protect\hyperlink{ref-Parker2006}{Parker, 2006}). As an example,
sizable areas of parameter space exists wherein the female optimum would
be to exhibit preference for (and receive) nuptial gifts, while the male
optimum is to not search for (and give) nuptial gifts (see Figure 2a,
Zone C). This will lead to mate encounters in which gift-less males will
benefit from mating, but females will not. In many systems, ecological
variables such as search time required to find a nuptial gift will
likely depend on prey abundance, which can vary substantially with time
in some species with nuptial gift-giving
(\protect\hyperlink{ref-Ghislandi2018}{Ghislandi \emph{et al.}, 2018}).
Since several ecological variables likely affect the value of these
thresholds, our results can be seen as providing some formalised
description of why nuptial gift-giving only occurs in some but not all
systems.

At first, the analytical model seems to suggests that nuptial gifts must
cause a very high fitness increase (approximately 25\%) before male
search and female choosiness is favoured by selection (Figure 2).
Similarly the IBM model seems to suggest that a fitness benefit of
approximately 50\% is required (see Figure 3). However, it is important
to recognise that these thresholds depend on multiple parameters. For
example, if female processing time (\(T_{\mathrm{f}}\)) is high, the
female threshold for choosiness with respect to \(\gamma\) drops such
that male search and female choosiness are favoured at lower \(\gamma\)
(see Supporting Information S2). If \(T_{\mathrm{f}}\) is sufficiently
high, then an initially rare gift-giving trait might be favoured by
selection even if the fitness benefit of a nuptial gift is low. The
effect that nuptial gifts have on fitness might vary across species, or
even populations. Effects on female fecundity have been estimated in
crickets, fireflies, butterflies, and spiders, but these estimates vary
considerably between species suggesting a large positive effect to no
effect at all (\protect\hyperlink{ref-Bergstrom2002}{Bergström \&
Wiklund, 2002}; \protect\hyperlink{ref-Rooney2002}{Rooney \& Lewis,
2002}; \protect\hyperlink{ref-Maxwell2018}{Maxwell \& Prokop, 2018};
\protect\hyperlink{ref-Gao2019}{Gao \emph{et al.}, 2019}).

We modelled the evolution of nuptial gift-giving using both a
mathematical model and an individual-based model. Our mathematical model
makes simple assumptions about the relationship between nuptial gift
search time (\(\alpha\)), conspecific encounter rate (\(R\)), female
processing time (\(T_{\mathrm{f}}\)), and the fitness increment of a
nuptial gift for offspring fitness (\(\gamma\)). It then derives the
threshold \(\gamma\) values above which males increase their fitness by
searching (\(T_{\mathrm{m}} > 0\)) for a nuptial gift
(\(\gamma_{\mathrm{m}}\)) and females increase their fitness by choosing
to reject males without gifts (\(\gamma_{\mathrm{f}}\)). In contrast,
our IBM models individuals over discrete time steps, and key processes
of nuptial gift acquisition, conspecific encounters, and female
processing are stochastic and varying among individuals. The
mathematical model and IBM make qualitatively identical predictions
(compare Figure 2b versus Figure 3), but differences between the two
models inevitably lead to quantitative differences for
\(\gamma_{\mathrm{m}}\) and \(\gamma_{\mathrm{f}}\) thresholds. For
example, \(\gamma_{\mathrm{f}}\) increased more rapidly with increasing
\(\alpha\) in the IBM compared to the analytical model. Some differences
are expected to occur due to stochastic effects inherent to IBMs (e.g.,
\protect\hyperlink{ref-Wilson2003}{Wilson \emph{et al.}, 2003}). Other
differences are more likely caused by more subtle assumptions between,
and limitations of, the two models. In general, the IBM did not do a
good job of controlling for conspecific interaction rate, making it
difficult to directly compare \(R\) between models. The IBM also allowed
for coevolution between male search and female choosiness (Figure 3),
which was not allowed in the analytical model (Figure 2). It was not our
goal to exactly recover the quantitative predictions of the analytical
model in our IBM. Future development of the IBM could further bridge the
gap between models while also developing new theory on how aspects of
the system such as explicit space, individual life history, or genetics
affect the evolution of nuptial gift-giving behaviour.

\hypertarget{nuptial-gift-giving-theory}{%
\subsection{Nuptial gift-giving
theory}\label{nuptial-gift-giving-theory}}

When modelling nuptial gift evolution, the challenge is to construct a
framework that captures the frequency-dependent selection between male
nuptial gift-giving and female choosiness for nuptial gifts, and we do
this using a time-in, time-out model. Recent studies have modelled some
frequency-dependent aspect of nuptial gift giving using evolutionary
game theory (\protect\hyperlink{ref-MaynardSmith1982}{Maynard Smith,
1982}; \protect\hyperlink{ref-Vincent2005}{Vincent \& Brown, 2005}). Two
such studies formulated a quantitative genetics model to study
evolutionarily stable nuptial gift sizes in populations where the female
propensity to re-mate was evolving
(\protect\hyperlink{ref-Kamimura2021}{Kamimura \emph{et al.}, 2021};
\protect\hyperlink{ref-Iwasa2022}{Iwasa \& Yamaguchi, 2022}). The
results obtained in these studies complement our results by giving
equilibrium solutions to the evolutionary stable nuptial gift size,
whereas we determine the general conditions under which nuptial
gift-giving will evolve as given by the inequalities we derive.

Other modelling frameworks have made general predictions about sexually
selected traits, and these predictions are not mutually exclusive to
those made by our model. For example, the good genes hypothesis predicts
that costly traits such as nuptial gift-giving can be favoured since
males enduring the cost of a nuptial gift signals to females that their
genes confer high fitness precisely because they can afford this cost
(\protect\hyperlink{ref-Kirkpatrick1996}{Kirkpatrick, 1996};
\protect\hyperlink{ref-Byers2006}{Byers \& Waits, 2006}; but see
\protect\hyperlink{ref-Fromhage2022}{Fromhage \& Henshaw, 2022}). In
other words, costly sexually selected traits are favoured because they
are indicators of overall genetic quality
(\protect\hyperlink{ref-Martinossi-Allibert2019}{Martinossi-Allibert
\emph{et al.}, 2019}). Because of this, nuptial gift-giving could be a
case of condition-dependence where engaging in nuptial gift-giving is
only favourable for male in good condition (e.g., males capable of
successful search (\protect\hyperlink{ref-MaynardSmith1982}{Maynard
Smith, 1982}; \protect\hyperlink{ref-Engqvist2015}{Engqvist \& Taborsky,
2015}; \protect\hyperlink{ref-Ghislandi2018}{Ghislandi \emph{et al.},
2018})). In general, our model demonstrates how nuptial gift-giving
initially evolves before other mechanisms, such as good gene effects,
become relevant.

A nuptial gift can also constitute a dishonest signal of good body
condition since worthless, deceptive nuptial gifts have evolved in
several systems (\protect\hyperlink{ref-LeBas2005}{LeBas \& Hockham,
2005}; \protect\hyperlink{ref-Ghislandi2014}{Ghislandi \emph{et al.},
2014}). This is also the case in \emph{P. mirabilis} where males will
wrap plant parts or an empty exoskeleton in silk, as opposed to an
arthropod prey, and use this as a nuptial gift
(\protect\hyperlink{ref-Albo2011}{Albo \emph{et al.}, 2011};
\protect\hyperlink{ref-Ghislandi2014}{Ghislandi \emph{et al.}, 2014}).
In such systems, worthless nuptial gifts have been shown to reduce the
likelihood that a male is rejected by a female compared to the case
where no nuptial gift is given. However, males offering worthless
nuptial gifts may be at a slight disadvantage in sperm competition since
worthless gifts result in a shorter copulation duration and hence less
sperm transfer (\protect\hyperlink{ref-Albo2013}{Albo \emph{et al.},
2013}; \protect\hyperlink{ref-Ghislandi2014}{Ghislandi \emph{et al.},
2014}). Worthless gifts should not result in any paternal care benefits
to the male since the offspring he may sire will not gain nutrition from
a worthless nuptial gift.

Given our modelling framework, worthless nuptial gifts may be expected
to evolve in cases where females are discriminating in favour of nuptial
gifts, but the cost of search time for a true nuptial gift is very high
such that selection will not favour male search. This scenario would
correspond to zone C of Figure 2 where the value of the nuptial gift
exceeds the female fitness threshold for choosiness to be favoured, but
due to high search time, selection will not favour male search for true
nuptial gifts. Our model also predicts the possibility of the opposite
scenario, in which males provide nuptial gifts, but females do not
exhibit preference for nuptial gifts (zone B of Figure 2).
Fascinatingly, an example of such system has been documented by a recent
study of the genus \emph{Trechaleoides}, which contains two species with
true nuptial gift-giving, but a lack of preference for nuptial gifts
among female (\protect\hyperlink{ref-Martinez2023}{Martínez Villar
\emph{et al.}, 2023}).

The main drivers of male nuptial gift-giving are thought to be indirect
fitness benefits and increased success in sperm competition, since
providing a nuptial gift can result in longer copulation duration which
is correlated with increased sperm transfer along with female cryptic
choice promoting males who provide nuptial gifts
(\protect\hyperlink{ref-Albo2011}{Albo \emph{et al.}, 2011},
\protect\hyperlink{ref-Albo2013}{2013}). However, nuptial gifts might
also function to modulate female aggression and prevent sexual
cannibalism (\protect\hyperlink{ref-Bilde2006}{Bilde \emph{et al.},
2006}). In some systems, such as \emph{P. mirabilis}, males have been
shown to reduce the risk of being cannibalised by the female after
mating when offering a nuptial gift, such that the nuptial gift may
result in a ``shield effect'', protecting the male
(\protect\hyperlink{ref-Toft2016}{Toft \& Albo, 2016}).

\hypertarget{empirical-implications}{%
\subsection{Empirical implications}\label{empirical-implications}}

The simulations parameterised with an experimentally estimated value of
\(\gamma\) showed evolution of nuptial gift searching in males and
choosiness for nuptial gifts in females. The model thus predicts that
\emph{P. mirabilis} living under conditions with the estimated fitness
value of nuptial gifts should exhibit both search for nuptial gifts and
choosiness for males with nuptial gifts, and this is what is observed in
empirical populations. Parameterising \(\gamma\) with data from
experimental studies may only yield a rough approximation of the true
\(\gamma\). This is because the estimated value of \(\gamma\) is based
on data from current populations (rather than ancestral populations,
which are being simulated), and because the literature is inconclusive
as to how much (if any) effect nuptial gifts have on female fitness
(\protect\hyperlink{ref-Maxwell2018}{Maxwell \& Prokop, 2018}). The
effect of nuptial gifts on female fecundity has been estimated in a
variety of system such as crickets, fireflies, butterflies and spiders,
but these estimates vary considerably between species suggesting a large
positive effect to no effect at all
(\protect\hyperlink{ref-Bergstrom2002}{Bergström \& Wiklund, 2002};
\protect\hyperlink{ref-Rooney2002}{Rooney \& Lewis, 2002};
\protect\hyperlink{ref-Maxwell2018}{Maxwell \& Prokop, 2018};
\protect\hyperlink{ref-Gao2019}{Gao \emph{et al.}, 2019}).

Our model assumes something akin to sequential polyandry. That is, a
system wherein female mating and reproduction with multiple males occurs
in sequence, rather than multiple matings occurring before reproduction.
In some systems with nuptial gift-giving, females have been documented
to mate multiple times before reproduction occurs, including the genus
of bark lice \emph{Neotrogla}
(\protect\hyperlink{ref-Kamimura2021}{Kamimura \emph{et al.}, 2021}),
and even our example system of \emph{P. mirabilis} where females will
sometimes engage in multiple mating before reproducing, especially if
starved because multiple mating may result in more nuptial gifts
(\protect\hyperlink{ref-Toft2015}{Toft \& Albo, 2015};
\protect\hyperlink{ref-Matzke2022}{Matzke \emph{et al.}, 2022}). It is
unclear what effect (if any) assuming non-sequential polyandry would
have on the threshold we derive. Under non-sequential polyandry, a
viable strategy for females might be to accept any male (with or without
gift) for fertilisation assurance, then exhibit a preference for nuptial
gifts. This might make choosiness less costly since it would entail less
of an opportunity cost to be choosy, and this could potentially make
female preference for nuptial gifts more likely to evolve. Our model
also assumes that males search in time-out, rather than contribute to
parental care, which is likely to be accurate for most systems but not
all. Expanding the model to explore these possibilities would be a
worthwhile goal for future research.

\hypertarget{conclusion}{%
\subsection{Conclusion}\label{conclusion}}

Overall, we found that a simple relationship between nuptial gift search
time and mate encounter rate yields a threshold that determines whether
selection will favour males that search for nuptial gifts. Similarly, we
found that the threshold determining whether females will be favoured to
reject males without nuptial gifts is also dependent on these variables,
along with offspring processing time. Together, these thresholds
describe the conditions under which nuptial gift-giving is expected to
evolve. The applications of these thresholds are numerous. They can be
used as a starting point for more complex or more system-specific models
of nuptial gift-giving evolution. They can also provide novel insight
into how populations can evolve to use worthless or token nuptial gifts.

\hypertarget{methods}{%
\section{Methods}\label{methods}}

\hypertarget{model}{%
\subsection{Model}\label{model}}

Here we first present more detail for the derivation of fitness
threshold values \(\gamma_{\mathrm{m}}\) and \(\gamma_{\mathrm{f}}\). We
then present IBM simulations (see Supporting Information S1 for full
details). Code for simulations is available on GitHub (see ``Data
availability'').

\hypertarget{derivation-of-fitness-thresholds}{%
\subsection{Derivation of fitness
thresholds}\label{derivation-of-fitness-thresholds}}

We use a time-in and time-out model in which females and males spend
some time searching for a mate (time-in) followed by a period of cool
down outside the mating pool (time-out; Figure 1).

After mating, females must spend some time processing offspring
(\(T_{\mathrm{f}}\)). Male time to replenish sperm is assumed to be
negligible, but males can spend time out of the mating pool
(\(T_{\mathrm{m}}\)) to search for a nuptial gift. When males return
from time-out, they encounter females with some probability that is a
function of the rate at which an individual encounters conspecifics
(\(R\)) and the sex ratio (\(\beta\); males/females). Mortality occurs
for females and males in (\(\mu_{\mathrm{i,f}}\),
\(\mu_{\mathrm{i,m}}\)) and out (\(\mu_{\mathrm{o,f}}\),
\(\mu_{\mathrm{o,m}}\)) of the mating pool. Following Kokko \& Ots
(\protect\hyperlink{ref-Kokko2006}{2006}), we assume
\(m_{\mathrm{i,f}} = m_{\mathrm{o,f}} = 1\) and
\(m_{\mathrm{i,m}} = m_{\mathrm{o,m}} = 1\). While this choice is
arbitrary, we conducted a sensitivity analysis which shows that the
mortality parameters have no influence on the propensity for male search
and female choice to evolve (Supporting Information S2). First, we
describe the fitness consequences of male search time for a nuptial
gift. We then describe the fitness consequences of female choice to
accept or reject males based on their provision of a nuptial gift.

\hypertarget{male-fitness}{%
\subsection{Male fitness}\label{male-fitness}}

During time-out, males have the opportunity to search for a nuptial
gift. Males can adopt one of two strategies; either search or do not
search for a nuptial gift. Males with the former strategy continue to
search until they find a nuptial gift, while males that do not search
will immediately re-enter the mating pool. In this case, time searching
for a nuptial gift will come at the cost of mating opportunities but
might increase offspring fitness. We therefore need to model the
expected length of time \(E[T_{\mathrm{m}}]\) spent outside of the
mating pool for males that search for nuptial gifts, which is simply
\(\alpha\). Note that we can integrate search time \(t\) over the rate
at which nuptial gifts are encountered (\(\exp(-1/\alpha)\)) to show
\(E[T_{\mathrm{m}}] = \alpha\),

\[E[T_{\mathrm{m}}] = \int_{0}^{\infty}e^{- \frac{1}{\alpha}t}dt = \alpha.\]

The rate at which a focal male that searches for a nuptial gift
(\(\mathrm{M_{G}}\)) increases his fitness is therefore the fitness of
his offspring \((1 + \gamma)\) divided by expected time spent searching
for a nuptial gift (\(\alpha\)) plus time spent in the mating pool,
\((\beta + 1)/R\) (recall that females produce \(\lambda\) offspring),

\[W_{\mathrm{M_{G}}} = \lambda \frac{1 + \gamma}{\alpha + \left( \frac{\beta + 1}{R} \right)}.\]

In contrast, a male that does not search for a nuptial gift
(\(\mathrm{M_{L}}\)) has offspring with lower fitness, but spends less
time outside of the mating pool,

\[W_{\mathrm{M_{L}}} = \lambda \frac{1}{\left(\frac{\beta+1}{R} \right)} = \lambda \frac{R}{\beta + 1}.\]

We can then determine the conditions for which
\(W_{\mathrm{M_{G}}} > W_{\mathrm{M_{L}}}\), isolating \(\gamma\) to
find how large of a fitness benefit must be provided by the nuptial gift
to make the search cost worthwhile, which simplifies to ineq. 2. When
this inequality holds, males are favoured to search until they find a
nuptial gift, which would result in an average search time of
\(\alpha\). When the male trait is continuous (i.e., males search for
time period \(T_{\mathrm{m}}\)), it can be shown that the same threshold
can be reached by evaluating the partial derivative of the male fitness
function (Supporting Information S3). Hence, the thresholds are
consistent under different assumptions concerning male searching
strategy. Selection will cause males to search for nuptial gifts if the
fitness increase to offspring exceeds the product of search time and
female encounter rate.

\hypertarget{female-fitness}{%
\subsection{Female fitness}\label{female-fitness}}

During time-out, females process offspring over a duration of
\(T_{\mathrm{f}}\) (we assume that \(T_{\mathrm{f}} > \alpha\), else
females are not the choosy sex). When females re-enter the mating pool,
they encounter males at a rate of \(R\beta/(\beta + 1)\). If a female
encounters a male with a nuptial gift, we assume that she will mate with
him. But if a female encounters a male with no nuptial gift, then she
might accept or reject the male. If she rejects the male, then she will
remain in the mating pool. The rate at which a female encounters a male
with a nuptial gift is,

\[R_{\mathrm{F_{G}}} = R \left(\frac{\beta}{\beta + 1}\right)\left(1 - e^{-\frac{1}{\alpha}T_{m}}\right).\]

We can similarly model the rate at which a female encounters a gift-less
male,

\[R_{\mathrm{F_{L}}} = R \left(\frac{\beta}{\beta + 1}\right)\left(e^{-\frac{1}{\alpha}T_{m}}\right).\]

Note that we can recover the rate at which a female encounters any male,

\[R \left(\frac{\beta}{\beta + 1}\right) = R \left(\frac{\beta}{\beta + 1}\right) \left(1 - e^{-\frac{1}{\alpha}T_{\mathrm{m}}}\right) + R \left(\frac{\beta}{\beta + 1}\right) \left(e^{-\frac{1}{\alpha}T_{\mathrm{m}}}\right).\]

If \(R_{\mathrm{F_{G}}}\) is sufficiently high and
\(R_{\mathrm{F_{L}}}\) is sufficiently low, then finding a male with a
gift will be easier than finding a male without one. Also, the expected
time spent in the mating pool before a focal female encounters a male
with a gift will be \(1/R_{\mathrm{F_{G}}}\), while the expected time
spent in the mating pool before a focal female encounters any male will
be \(1 / (R_{\mathrm{F_{G}}} + R_{\mathrm{F_{L}}})\). Finally, the rates
at which a female encounters males with and without a gift,
\(R_{\mathrm{F_{G}}}\) and \(R_{\mathrm{F_{L}}}\), are different from
the probabilities that a male encounter has or does not have a gift. The
rate of encounter is no longer relevant in this case because we are
assuming that an encounter has occurred. Hence, the probability of an
encountered male having a gift is simply,

\[P(G) = \frac{1 - e^{-\frac{1}{\alpha}T_{\mathrm{m}}}}{\left(1 - e^{-\frac{1}{\alpha}T_{\mathrm{m}}}\right) + e^{-\frac{1}{\alpha}T_{\mathrm{m}}}} = 1 - e^{-\frac{1}{\alpha}T_{\mathrm{m}}}.\]

Similarly, the probability of an encountered male not having a gift is
simply,

\[P(L) = e^{-\frac{1}{\alpha}T_{\mathrm{m}}}.\]

The rate at which a female increases her fitness by being choosy and
mating only when she encounters a male with a gift is,

\begin{equation}
W_{\mathrm{F_{G}}} = \lambda \frac{1 + \gamma}{T_{\mathrm{f}}} + \frac{1}{\mathrm{F_{G}}}.
\end{equation}

The top of the right-hand side of Eq. 6 gives the fitness increase, and
the bottom gives the total time it takes to obtain this fitness. The
\(R_{\mathrm{F_{G}}}\) is inverted because it represents the expected
time to encountering a male with a gift. We can expand Eq. 6,

\[W_{\mathrm{F_{G}}} = \lambda \frac{1 + \gamma}{T_{\mathrm{f}} + \frac{1}{R \left(\frac{\beta}{\beta + 1}\right)\left(1 - e^{-\frac{1}{\alpha}T_{\mathrm{m}}}\right)}}.\]

If the focal female is not choosy and accepts the first male that she
encounters, then the rate at which she increases her fitness is,

\[W_{\mathrm{F_{G,L}}} = \lambda \frac{\left(1 + \gamma\right)\left(1 - e^{-\frac{1}{\alpha}T_{\mathrm{m}}}\right) + e^{-\frac{1}{\alpha}T_{\mathrm{m}}}}{T_{f} + \frac{1}{R \left(\frac{\beta}{\beta + 1}\right)}}.\]

We then evaluate the conditions under which
\(W_{\mathrm{F_{G}}} > W_{\mathrm{F_{G,L}}}\). We isolate \(\gamma\) to
determine how much offspring fitness must be increased to make
choosiness beneficial (\(\gamma_{\mathrm{f}}\)),

\[\gamma_{\mathrm{f}} > \frac{1 + \frac{1}{\beta}}{D T_{f} \left(1 - e^{-1\frac{1}{\alpha}T_{m}}\right)}.\]

The above can be expressed as Eq. 7 below,

\begin{equation}
\gamma_{\mathrm{f}} > \frac{1}{T_{\mathrm{f}}} R\left(\frac{\beta}{\beta + 1}\right) \left(1 - e^{-1\frac{1}{\alpha}T_{\mathrm{m}}}\right).
\end{equation}

Note that that the expression
\(R\left(\frac{\beta}{\beta + 1}\right) \left(1 - e^{-1\frac{1}{\alpha}T_{\mathrm{m}}}\right)\)
defines the rate at which a female in the mating pool encounters males
with nuptial gifts. Hence, female choosiness is ultimately determined by
time spent out of the mating pool to process offspring
(\(T_{\mathrm{f}}\)) and the rate at which a female in the mating pool
encounters males with nuptial gifts.

\hypertarget{operational-sex-ratio}{%
\subsection{Operational sex ratio}\label{operational-sex-ratio}}

We assume that the sex ratio is equal upon maturation. Given this, Kokko
\& Monaghan (\protect\hyperlink{ref-Kokko2001}{2001}) show that the
operational sex ratio depends on the probability of finding an
individual in time-in,

\begin{equation}
\beta = \frac{\int_{t = 0}^{\infty} P_{IM}(t)dt}{\int_{t = 0}^{\infty} P_{IF}(t)dt}.
\end{equation}

In Eq. 8, \(P_{IM}(t)\) and \(P_{IF}(t)\) are the probabilities of
finding a male and female in time-in, respectively. There is no closed
form solution to the operational sex ratio, so we used recursion to
calculate \(\beta\) values for a given \(T_{\mathrm{f}}\),
\(T_{\mathrm{m}}\), and \(R\) (see Supporting Information S4),

\hypertarget{individual-based-model}{%
\subsection{Individual-based model}\label{individual-based-model}}

We formulate an individual-based simulation model to test whether the
predictions made by the analytical time-in time-out model are
qualitatively the same under a similar simulation model. We use the
individual-based model to test whether the prediction hold in finite
populations. The IBM was written in C. All details of the
initialisation, parameterisation and the specific simulations which we
run, are described in Supporting Information S1.

We use available experimental data on the effect of nuptial gifts on
female offspring production to estimate the key parameter \(\gamma\)
(fitness increment from nuptial gift) and conducted a series of
simulations where \(\gamma\) was parameterised using this estimated
value. Details of the estimation of \(\gamma\) are described in
Supporting Information S5. Separate evolution of male search and female
choice (i.e., without co-evolution) is simulated in Supporting
Information S6.

\textbf{Author contributions:} APC and ABD conceived the study. ABD
constructed the modelling framework with input from APC. APC wrote the
paper with input from ABD and ABD wrote the IBM model. TB and GB
provided substantial comments on previous drafts and final text.

\textbf{Acknowledgements:} Anders P. Charmouh was supported by the
University of Aberdeen. A. Bradley Duthie was supported by the
University of Stirling. Trine Bilde was supported by The Danish Council
for Independent Research grant number 4002-00328B. We thank Maria Albo
for making available data from a previous study which was useful for
parameter estimating.

\textbf{Data availability:} The simulation software was implemented in C
and the full source code is available at
\url{https://github.com/bradduthie/nuptial_gift_evolution}.

\textbf{Competing interests:} The authors declare no competing
interests.

\textbf{References}

\hypertarget{refs}{}
\begin{CSLReferences}{0}{0}
\leavevmode\vadjust pre{\hypertarget{ref-Albo2013}{}}%
Albo, M.J., Bilde, T. \& Uhl, G. (2013)
\href{https://doi.org/10.1098/rspb.2013.1735}{{Sperm storage mediated by
cryptic female choice for nuptial gifts}}. \emph{Proceedings of the
Royal Society B: Biological Sciences}, \textbf{280}.

\leavevmode\vadjust pre{\hypertarget{ref-Albo2019}{}}%
Albo, M.J., Franco-Trecu, V., Wojciechowski, F.J., Toft, S. \& Bilde, T.
(2019) \href{https://doi.org/10.1093/beheco/arz040}{{Maintenance of
deceptive gifts in a natural spider population: Ecological and
demographic factors}}. \emph{Behavioral Ecology}, \textbf{30},
993--1000.

\leavevmode\vadjust pre{\hypertarget{ref-Albo2014}{}}%
Albo, M.J., Toft, S. \& Bilde, T. (2014) {Sexual selection, ecology, and
evolution of nuptial gifts in spiders}. In \emph{Sexual selection:
Perspectives and models from the neotropics} (ed. by Macedo, R.H. \&
Machado, G.). Academic Press, London, pp. 183--200.

\leavevmode\vadjust pre{\hypertarget{ref-Albo2011}{}}%
Albo, M.J., Winther, G., Tuni, C., Toft, S. \& Bilde, T. (2011)
\href{https://doi.org/10.1186/1471-2148-11-329}{{Worthless donations:
Male deception and female counter play in a nuptial gift-giving
spider}}. \emph{BMC Evolutionary Biology}, \textbf{11}.

\leavevmode\vadjust pre{\hypertarget{ref-Arnqvist2005a}{}}%
Arnqvist, G. \& Rowe, L. (2005) \emph{{Sexual Conflict}}. Princeton
University Press, Princeton, New Jersey.

\leavevmode\vadjust pre{\hypertarget{ref-Bergstrom2002}{}}%
Bergström, J. \& Wiklund, C. (2002) Effects of size and nuptial gifts on
butterfly reproduction: Can females compensate for a smaller size
through male-derived nutrients? \emph{Behavioral Ecology and
Sociobiology}, \textbf{52}, 296--302.

\leavevmode\vadjust pre{\hypertarget{ref-Bilde2006}{}}%
Bilde, T., Tuni, C., Elsayed, R., Pekár, S. \& Toft, S. (2006)
\href{https://doi.org/10.1098/rsbl.2005.0392}{{Death feigning in the
face of sexual cannibalism}}. \emph{Biology Letters}, \textbf{2},
23--25.

\leavevmode\vadjust pre{\hypertarget{ref-Bristowe1926}{}}%
Bristowe, W. \& Locket, G.H. (1926) {The courtship of British lycosid
spiders, and its probable significance}. \emph{Proceedings of the
Zoological Society of London}, \textbf{96}, 317--347.

\leavevmode\vadjust pre{\hypertarget{ref-Byers2006}{}}%
Byers, J.A. \& Waits, L. (2006)
\href{https://doi.org/10.1073/pnas.0608184103}{{Good genes sexual
selection in nature}}. \emph{Proceedings of the National Academy of
Sciences}, \textbf{103}, 16343--16345.

\leavevmode\vadjust pre{\hypertarget{ref-Clutton-Brock1992}{}}%
Clutton-Brock, T.H. \& Parker, G.A. (1992)
\href{https://doi.org/10.1086/417793}{{Potential reproductive rates and
the operation of sexual selection}}. \emph{Quarterly Review of Biology},
\textbf{67}, 437--456.

\leavevmode\vadjust pre{\hypertarget{ref-Engqvist2015}{}}%
Engqvist, L. \& Taborsky, M. (2015)
\href{https://doi.org/10.1098/rspb.2015.2945}{{The evolution of genetic
and conditional alternative reproductive tactics}}. \emph{Proceedings of
the Royal Society B: Biological Sciences}, \textbf{283}.

\leavevmode\vadjust pre{\hypertarget{ref-Fromhage2022}{}}%
Fromhage, L. \& Henshaw, J.M. (2022) The balance model of honest sexual
signaling. \emph{Evolution}, \textbf{76}, 445--454.

\leavevmode\vadjust pre{\hypertarget{ref-Gao2019}{}}%
Gao, Q., Turnell, B.R., Hua, B. \& Shaw, K.L. (2019) The effect of
nuptial gift number on fertilization success in a hawaiian swordtail
cricket. \emph{Behavioral Ecology and Sociobiology}, \textbf{73}, 1--9.

\leavevmode\vadjust pre{\hypertarget{ref-Ghislandi2014}{}}%
Ghislandi, P.G., Albo, M.J., Tuni, C. \& Bilden, T. (2014)
\href{https://www.ptonline.com/articles/how-to-get-better-mfi-results}{{Evolution
of deceit by worthless donations in a nuptial gift-giving spider}}.
\emph{Current Zoology}, \textbf{60}, 43--51.

\leavevmode\vadjust pre{\hypertarget{ref-Ghislandi2018}{}}%
Ghislandi, P.G., Pekár, S., Matzke, M., Schulte-Döinghaus, S., Bilde, T.
\& Tuni, C. (2018) \href{https://doi.org/10.1111/jeb.13284}{{Resource
availability, mating opportunity and sexual selection intensity
influence the expression of male alternative reproductive tactics}}.
\emph{Journal of Evolutionary Biology}, \textbf{31}, 1035--1046.

\leavevmode\vadjust pre{\hypertarget{ref-Iwasa2022}{}}%
Iwasa, Y. \& Yamaguchi, S. (2022)
\href{https://doi.org/10.1016/j.jtbi.2021.110939}{{Evolution of male
nuptial gift and female remating: A quantitative genetic model}}.
\emph{Journal of Theoretical Biology}, \textbf{533}, 110939.

\leavevmode\vadjust pre{\hypertarget{ref-Kamimura2021}{}}%
Kamimura, Y., Yoshizawa, K., Lienhard, C., Ferreira, R.L. \& Abe, J.
(2021) \href{https://doi.org/10.1186/s12862-021-01901-x}{{Evolution of
nuptial gifts and its coevolutionary dynamics with male-like persistence
traits of females for multiple mating}}. \emph{BMC Ecology and
Evolution}, \textbf{21}, 1--14.

\leavevmode\vadjust pre{\hypertarget{ref-Kirkpatrick1996}{}}%
Kirkpatrick, M. (1996) Good genes and direct selection in the evolution
of mating preferences. \emph{Evolution}, \textbf{50}, 2125--2140.

\leavevmode\vadjust pre{\hypertarget{ref-Kokko2001}{}}%
Kokko, H. \& Monaghan, P. (2001)
\href{https://doi.org/10.1046/j.1461-0248.2001.00212.x}{{Predicting the
direction of sexual selection}}. \emph{Ecology Letters}, \textbf{4},
159--165.

\leavevmode\vadjust pre{\hypertarget{ref-Kokko2006}{}}%
Kokko, H. \& Ots, I. (2006) {When not to avoid inbreeding}.
\emph{Evolution}, \textbf{60}, 467--475.

\leavevmode\vadjust pre{\hypertarget{ref-LeBas2005}{}}%
LeBas, N.R. \& Hockham, L.R. (2005) \href{https://doi.org/10.1016/j}{{An
invasion of cheats: the evolution of worthless nuptial gifts}}.
\emph{Current Biology}, \textbf{15}, 64--67.

\leavevmode\vadjust pre{\hypertarget{ref-Lewis2014}{}}%
Lewis, S.M., Vahed, K., Koene, J.M., Engqvist, L., Bussière, L.F.,
Perry, J.C., \emph{et al.} (2014)
\href{https://doi.org/10.1098/rsbl.2014.0336}{{Emerging issues in the
evolution of animal nuptial gifts}}. \emph{Biology Letters},
\textbf{10}, 3--5.

\leavevmode\vadjust pre{\hypertarget{ref-Lewis2012}{}}%
Lewis, S. \& South, A. (2012)
\emph{\href{https://doi.org/10.1016/B978-0-12-394288-3.00002-2}{{The
Evolution of Animal Nuptial Gifts}}}.

\leavevmode\vadjust pre{\hypertarget{ref-Lubin2007}{}}%
Lubin, Y. \& Bilde, T. (2007)
\href{https://doi.org/10.1016/S0065-3454(07)37003-4}{{The evolution of
sociality in spiders}}. \emph{Advances in the Study of Behavior},
\textbf{37}, 83--145.

\leavevmode\vadjust pre{\hypertarget{ref-Martinez2023}{}}%
Martínez Villar, M., Germil, M., Pavón-Peláez, C., Tomasco, I., Bilde,
T., Toft, S., \emph{et al.} (2023) Lack of female preference for nuptial
gifts may have led to loss of the male sexual trait. \emph{Evolutionary
Biology}, \textbf{50}, 318--331.

\leavevmode\vadjust pre{\hypertarget{ref-Martinossi-Allibert2019}{}}%
Martinossi-Allibert, I., Rueffler, C., Arnqvist, G. \& Berger, D. (2019)
\href{https://doi.org/10.1098/rspb.2018.2313}{The efficacy of good genes
sexual selection under environmental change}. \emph{Proceedings of the
Royal Society B: Biological Sciences}, \textbf{286}.

\leavevmode\vadjust pre{\hypertarget{ref-Matzke2022}{}}%
Matzke, M., Toft, S., Bechsgaard, J., Pold Vilstrup, A., Uhl, G.,
Künzel, S., \emph{et al.} (2022) Sperm competition intensity affects
sperm precedence patterns in a polyandrous gift-giving spider.
\emph{Molecular Ecology}, \textbf{31}, 2435--2452.

\leavevmode\vadjust pre{\hypertarget{ref-Maxwell2018}{}}%
Maxwell, M.R. \& Prokop, P. (2018) Fitness effects of nuptial gifts in
the spider \emph{pisaura mirabilis}: Examination under an alternative
feeding regime. \emph{The Journal of Arachnology}, \textbf{46},
404--412.

\leavevmode\vadjust pre{\hypertarget{ref-MaynardSmith1982}{}}%
Maynard Smith, J. (1982)
\emph{\href{https://www.ncbi.nlm.nih.gov/pubmed/13761767}{{Evolution and
the theory of games}}}. Cambridge University Press, Cambridge.

\leavevmode\vadjust pre{\hypertarget{ref-Oliveira2008}{}}%
Oliveira, R.F., Taborsky, M. \& Brockmann, H.J. (Eds.). (2008)
\emph{{Alternative reproductive tactics: an integrative approach}}.
Cambridge University Press, Cambridge.

\leavevmode\vadjust pre{\hypertarget{ref-Parker2006}{}}%
Parker, G.A. (2006) \href{https://doi.org/10.1098/rstb.2005.1785}{Sexual
conflict over mating and fertilization: An overview}.
\emph{Philosophical Transactions of the Royal Society B}, \textbf{361},
235--259.

\leavevmode\vadjust pre{\hypertarget{ref-Rooney2002}{}}%
Rooney, J. \& Lewis, S.M. (2002) Fitness advantage from nuptial gifts in
female fireflies. \emph{Ecological Entomology}, \textbf{27}, 373--377.

\leavevmode\vadjust pre{\hypertarget{ref-Toft2015}{}}%
Toft, S. \& Albo, M.J. (2015) Optimal numbers of matings: The
conditional balance between benefits and costs of mating for females of
a nuptial gift-giving spider. \emph{Journal of Evolutionary Biology},
\textbf{28}, 457--467.

\leavevmode\vadjust pre{\hypertarget{ref-Toft2016}{}}%
Toft, S. \& Albo, M.J. (2016)
\href{https://doi.org/10.1098/rsbl.2015.1082}{{The shield effect:
Nuptial gifts protect males against pre-copulatory sexual cannibalism}}.
\emph{Biology Letters}, \textbf{12}.

\leavevmode\vadjust pre{\hypertarget{ref-Tuni2013a}{}}%
Tuni, C., Albo, M.J. \& Bilde, T. (2013)
\href{https://doi.org/10.1111/jeb.12137}{{Polyandrous females acquire
indirect benefits in a nuptial feeding species}}. \emph{Journal of
Evolutionary Biology}, \textbf{26}, 1307--1316.

\leavevmode\vadjust pre{\hypertarget{ref-Vincent2005}{}}%
Vincent, T.L. \& Brown, J.S. (2005) \emph{{Evolutionary Game Theory,
Natural Selection, and Darwinian Dynamics}}. Cambridge University Press,
Cambridge.

\leavevmode\vadjust pre{\hypertarget{ref-Wilson2003}{}}%
Wilson, W.G., Morris, W.F. \& Bronstein, J.L. (2003) {Coexistence of
mutualists and exploiters on spatial landscapes}. \emph{Ecological
Monographs}, \textbf{73}, 397--413.

\end{CSLReferences}

\end{document}